Sequential rank agreement methods for comparison of ranked lists

### 2016-04-08 latex/paper.tex
Intro, forfatterliste, bibtex

\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[]{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[]{color}
\usepackage{subcaption}
\usepackage{natbib}
\usepackage{xspace}
\usepackage{calc}
\usepackage[noae]{Sweave}
\usepackage{authblk}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{enumitem}
%\usepackage{geometry}
%\geometry{a4paper, portrait, margin=1in}
%
% New commands
%
\newcommand{\ie}{\emph{i.e.}\@\xspace}
\newcommand{\eg}{\emph{e.g.}\@\xspace}

@@

\title{Sequential rank agreement methods for comparison of ranked lists}
\author{Claus Thorn Ekstrøm}
\author{Thomas Alexander Gerds}
\author{Andreas Kryger Jensen}
\author{Kasper Brink-Jensen}
\affil{Biostatistics, University of Copenhagen}
\renewcommand{\topfraction}{.85}
\renewcommand{\bottomfraction}{.7}
\renewcommand{\textfraction}{.15}
\renewcommand{\floatpagefraction}{.66}
\renewcommand{\dbltopfraction}{.66}
\renewcommand{\dblfloatpagefraction}{.66}
\setcounter{topnumber}{9}
\setcounter{bottomnumber}{9}
\setcounter{totalnumber}{20}
\setcounter{dbltopnumber}{9}

@@

In this article we introduce sequential rank agreement for measuring
agreement among (partially) ranked lists.  The general idea is to
define agreement based on the sequence of ranks from the first $d$
elements in each list. As agreement metric we adapt the limits of
agreement known from agreement between quantitative variables
\citep{alt:bland:1983,Carstensen2010}. Our proposed approach allows us
to compare more than two lists simultaneously, it provides a dynamic
measure of agreement as a function of the depth in the lists, places
high weight on the top of the list, accommodates censored/incomplete
lists of varying lengths, and has a natural interpretation that
directly relates to the ranks. Graphical illustration of sequential
rank agreement potentially allows us to infer a change-point, i.e., a
list depth where a change in the agreement of the lists occur but we
also provide randomization-based graphical tools to compare the
observed rank agreement to the expected agreement found in
non-informative data.  In this sense our approach is a combination and
generalization of some of the ideas of \citet{Carterette2009} and
\citet{Boulesteix2009}. The former compares two rankings based on the
distance between them as measured by a multivariate Gaussian
distribution and the latter presents an overview of approaches for
aggregation of ranked lists including bootstrap and leave-one-out
jackknife approaches.

The manuscript is organized as follows: In the next section we define
sequential rank agreement for multiple ranked lists and discuss how to
handle incomplete/censored/partial lists. In section 3 we present and
discuss approaches to evaluate the results obtained from sequential
rank agreement. Finally we apply the sequential rank agreement to two
Danish ovarian cancer studies before we discuss the findings along
with possible extensions. The approaches presented in this manuscript
are available in the R package \texttt{SuperRanker}.

@@

% latex table generated in R 3.2.2 by xtable 1.7-4 package
% Thu Apr  7 20:04:21 2016
\begin{table}[tb]
\centering
\caption{Top 10 list of ranked results from the Golub data. Numbers indicate the predictor/gene for the given ranking and method} 
\label{tab1}
\begin{tabular}{rrrrr}
  \hline
Ranking & T & LogReg & ElasticNet & MIC \\ 
  \hline
1 & 2124 & 2124 & 829 & 378 \\ 
  2 & 896 & 896 & 2124 & 829 \\ 
  3 & 2600 & 829 & 2198 & 896 \\ 
  4 & 766 & 394 & 1665 & 1037 \\ 
  5 & 829 & 766 & 1042 & 2124 \\ 
  6 & 2851 & 2670 & 808 & 808 \\ 
  7 & 703 & 2939 & 1920 & 108 \\ 
  8 & 2386 & 2386 & 849 & 515 \\ 
  9 & 2645 & 1834 & 523 & 2670 \\ 
  10 & 2002 & 378 & 937 & 2600 \\ 
   \hline
\end{tabular}
\end{table}

>>>>

\documentclass[12pt,a4paper]{article}
\pdfoutput=1
\usepackage[utf8]{inputenc}
\usepackage[]{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[]{color}
\usepackage{subcaption}
\usepackage{natbib}
\usepackage{xspace}
\usepackage{calc}
\usepackage[noae]{Sweave}
\usepackage{authblk}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{enumitem}
%\usepackage{geometry}
%\geometry{a4paper, portrait, margin=1in}
%
% New commands
%
\newcommand{\ie}{\emph{i.e.}\@\xspace}
\newcommand{\eg}{\emph{e.g.}\@\xspace}

@@

\title{Sequential rank agreement methods for comparison of ranked lists}
\author{Claus Thorn Ekstrøm}
\author{Thomas Alexander Gerds}
\author{Andreas Kryger Jensen}
%\author{Kasper Brink-Jensen}
\affil{Biostatistics, University of Copenhagen}
\renewcommand{\topfraction}{.85}
\renewcommand{\bottomfraction}{.7}
\renewcommand{\textfraction}{.15}
\renewcommand{\floatpagefraction}{.66}
\renewcommand{\dbltopfraction}{.66}
\renewcommand{\dblfloatpagefraction}{.66}
\setcounter{topnumber}{9}
\setcounter{bottomnumber}{9}
\setcounter{totalnumber}{20}
\setcounter{dbltopnumber}{9}

@@

Very recently, \citet{hall:schi:2012} proposed a method for comparing
pairwise rankings and derived the asymptotic distribution of the
endpoint where the two ranked lists no longer are in agreement. Their
approach was based on anchoring one of the two lists and subsequently
generating a sequence of 0s and 1s depending on whether the ranks in
the second list was close to the rank from the anchored
list. \citet{samp:verd:2013} followed up on this idea for pairwise
comparison of lists but used penalties based on a truncated geometric
probability instead of a 0-1 process and they evaluated the
distribution of the endpoint of agreement by computational approaches.
The asymptotic distribution in the \citet{hall:schi:2012} paper is
based on letting the number of \emph{lists} increase to infinity which
is a situation that is only relevant in special cases, and the
simulation-based null distribution approach of \citet{samp:verd:2013}
is guaranteed to work also for small samples.

In this article we introduce sequential rank agreement for measuring
agreement among (partially) ranked lists.  The general idea is to
define agreement based on the sequence of ranks from the first $d$
items in each list. As agreement metric we adapt the limits of
agreement known from agreement between quantitative variables
\citep{alt:bland:1983,Carstensen2010} but any measure for agreement
could be essentially be used. Our proposed approach allows us to
compare more than two lists simultaneously, it provides a dynamic
measure of agreement as a function of the depth in the lists, places
high weight on items at the top of the list, accommodates partially
observed lists of varying lengths, and has a natural interpretation
that directly relates to the ranks.  Graphical illustration of
sequential rank agreement potentially allows us to infer a
change-point, i.e., a list depth where a change in the agreement of
the lists occur but we also provide randomization-based graphical
tools to compare the observed rank agreement to the expected agreement
found in non-informative data.  In this sense our approach is a
combination and generalization of some of the ideas of
\citet{Carterette2009} and \citet{Boulesteix2009}. The former compares
two rankings based on the distance between them as measured by a
multivariate Gaussian distribution and the latter presents an overview
of approaches for aggregation of ranked lists including bootstrap and
leave-one-out jackknife approaches. We show the asymptotic
distribution of the endpoint of agreement and discuss how to infer the
distribution in small-sample situations using computational
methods. This enables us to make inferences about the endpoint of
agreement even for situations where there is no actual changepoint
present but just a gradual decline in agreement.

The manuscript is organized as follows: In the next section we define
sequential rank agreement for multiple ranked lists and discuss how to
handle incomplete/censored/partial lists. In section 3 we present and
discuss approaches to evaluate the results obtained from sequential
rank agreement. Finally we apply the sequential rank agreement to two
Danish ovarian cancer studies before we discuss the findings along
with possible extensions. The approaches presented in this manuscript
are available in the R package \texttt{SuperRanker}. [XXX check dette
passer, submit pakke]

@@

% latex table generated in R 3.2.3 by xtable 1.8-2 package
% Fri Apr  8 10:22:58 2016
\begin{table}[tb]
\centering
\caption{Top 10 list of ranked results from the Golub data. Numbers indicate the predictor/gene for the given ranking and method} 
\label{tab1}
\begin{tabular}{rrrrr}
  \hline
Ranking & T & LogReg & ElasticNet & MIC \\ 
  \hline
1 & 2124 & 2124 & 829 & 378 \\ 
  2 & 896 & 896 & 2124 & 829 \\ 
  3 & 2600 & 829 & 2198 & 896 \\ 
  4 & 766 & 394 & 1665 & 1037 \\ 
  5 & 829 & 766 & 1920 & 2124 \\ 
  6 & 2851 & 2670 & 1042 & 808 \\ 
  7 & 703 & 2939 & 808 & 108 \\ 
  8 & 2386 & 2386 & 849 & 515 \\ 
  9 & 2645 & 1834 & 937 & 2670 \\ 
  10 & 2002 & 378 & 1995 & 2600 \\ 
   \hline
\end{tabular}
\end{table}
